---
title: "Untitled"
author: "Jaiden Goerlitz"
date: "2023-03-19"
output: word_document
---

```{r setup, include=FALSE}
## Loading packages 
library(tidyverse)
library(readr)
library(plyr)
library(dplyr)
library(ggplot2)
```

Explatory Data Analysis

First, we will load the data.

```{r}
data1 <- read.csv("C:\\Users\\19goe\\Documents\\Spring 2023\\Stat 482\\Stat-482-Project\\Data\\SpotifyAudioFeaturesApril2019.csv")
view(data1)
```

Next, we want to see the distribution of poularity for the entire dataset. 

```{r}
popularity <- data1$popularity
hist(popularity)
```
First, since the distribution of the popularity in the data is so skewed we will create new dataframes. In the first dataframe we will randomly sample 100 songs for each popularity range of: 0-10, 11-20, 21-30, 31-40, 41-50, 51-60, 61-70, 71-80, 81-90, 91-100. 

```{r}

# Create an empty data frame to store the sampled data
sampled_data <- data.frame()

# Loop through each popularity range and add a random sample of 100 songs to the new data frame
for (i in 0:9) {
  # Filter the original data frame for songs with popularity score within the current range
  filtered_data <- data1[data1$popularity >= i*10 & data1$popularity < (i+1)*10,]
  
  # Check if the number of songs in the filtered data is at least 100
  if (nrow(filtered_data) >= 100) {
    # Take a random sample of 100 songs from the filtered data
    sampled_songs <- filtered_data[sample(nrow(filtered_data), 100),]
    
    # Add the sampled songs to the new data frame
    sampled_data <- rbind(sampled_data, sampled_songs)
  }
  else {
    # If there are fewer than 100 songs in the filtered data, add all of them to the new data frame
    sampled_data <- rbind(sampled_data, filtered_data)
  }
}

# Now you can use the "sampled_data" data frame for further analysis

```


```{r}
view(sampled_data)
```

```{r}
vars_to_plot <- setdiff(names(sampled_data), c("acousticness", "danceability", "duration_ms", "energy", "instrumentalness", "key", "liveness", "loudness", "mode", "speechiness", "tempo", "time_signature", "valence"))

```

```{r}
library(corrplot)

# Assume your sampled data frame is called "sampled_data"
# Select the columns to include in the correlation plot
columns <- c("popularity", "acousticness", "danceability", "duration_ms", "energy", "instrumentalness", "key", "liveness", "loudness", "mode", "speechiness", "tempo", "time_signature", "valence")

# Calculate the correlation matrix
corr_matrix <- cor(sampled_data[columns])

# Plot the correlation matrix
corrplot(corr_matrix, method = "circle", type = "upper", order = "hclust", col = colorRampPalette(c("#3366FF", "#FFFFFF", "#FF3333"))(200), addrect = 4)

```

We also tried several other sampling methods. The distribution of a songs popularity is so much higher for the score 0-10. We found the proportions of each popularity score range and based our sample sizes for each range on those proportions. They are follows:    

We decided for this dataframe that songs that are "popular" will have a score greater than or equal to 60. We will keep all of the the songs that have a score greater than 60 and randomly sample score ranges 0-9, 10-19, 20-29, 30-39, 40-49, 50-69 based on the proportions of each range from the original dataset. 

```{r}
# Assume your Spotify data is stored in a data frame called "spotify_data"

# Define the popularity score ranges
popularity_ranges <- seq(0, 100, by = 10)

# Group the popularity scores into bins
popularity_bins <- cut(data1$popularity, breaks = popularity_ranges, right = FALSE)

# Calculate the proportion of songs in each bin
popularity_proportions <- prop.table(table(popularity_bins))

# Print the results
cat("Popularity range\tProportion\n")
cat("------------------------------\n")
for (i in 1:length(popularity_proportions)) {
  cat(paste0("[", popularity_ranges[i], ", ", popularity_ranges[i+1]), ")\t\t", round(popularity_proportions[i] * 100, 2), "%\n")
}

```


```{r}
# Define the sample sizes for each popularity range
sample_sizes <- c(292, 170, 170, 138, 98, 77, 0, 0, 0, 0)

# Create an empty data frame to store the sampled data
sampled_data2 <- data.frame()

# Loop through each popularity range and sample the songs
for (i in 1:length(sample_sizes)) {
  # Extract the songs with the current popularity range
  current_songs <- data1[data1$popularity >= ((i-1)*10) & data1$popularity < (i*10), ]
  
  # If the current popularity range has a sample size of 0, just append the current_songs to sampled_data
  if (sample_sizes[i] == 0) {
    sampled_data2 <- rbind(sampled_data2, current_songs)
  } else {
    # Otherwise, randomly sample the specified number of songs and append them to sampled_data
    current_sample <- current_songs[sample(nrow(current_songs), sample_sizes[i]), ]
    sampled_data2 <- rbind(sampled_data2, current_sample)
  }
}

# Print the number of songs in each popularity range for the original data and the sampled data
cat("Original data:\n")
cat(table(cut(data1$popularity, breaks = seq(0, 100, by = 10), right = FALSE)), "\n")
cat("Sampled data:\n")
cat(table(cut(sampled_data2$popularity, breaks = seq(0, 100, by = 10), right = FALSE)), "\n")

```
```{r}
view(sampled_data2)
```

```{r}
library(corrplot)

# Assume your sampled data frame is called "sampled_data"
# Select the columns to include in the correlation plot
columns <- c("popularity", "acousticness", "danceability", "duration_ms", "energy", "instrumentalness", "key", "liveness", "loudness", "mode", "speechiness", "tempo", "time_signature", "valence")

# Calculate the correlation matrix
corr_matrix <- cor(sampled_data2[columns])

# Plot the correlation matrix
corrplot(corr_matrix, method = "circle", type = "upper", order = "hclust", col = colorRampPalette(c("#3366FF", "#FFFFFF", "#FF3333"))(200), addrect = 4)
```

```{r}
# Load the corrplot package
library(corrplot)

# Compute the correlation matrix for the features of interest
features <- c("acousticness", "danceability", "duration_ms", "energy", "instrumentalness", "key", "liveness", "loudness", "mode", "speechiness", "tempo", "time_signature", "valence", "popularity")
cor_matrix <- cor(sampled_data2[, features])

# Create a correlation plot with the "ellipse" method for highlighting high correlations
corrplot(cor_matrix, method = "ellipse", type = "upper", tl.col = "black", tl.srt = 45)

```

